Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code Verification

### 2015-03-04 papers/wdmerger_I/paper.tex
Correction to previous commit: three levels of refinement == 6.25 km.
First pass at a performance section

We desire to specify the mass of the white dwarf, as well as its
temperature and composition. To start the integration off, we
therefore need to guess at a central density.  We then do a secant
iteration over the entire integration procedure to find the central
density needed to yield the desired total mass.  The grid spacing is
$\Delta x = 25\ \text{km}$. We chose this value because no simulation
we perform is likely to exceed this grid resolution inside the stars 
themselves; for our normal domain size (see below), this corresponds to 
three jumps in refinement by a factor of four. We find that for low 
resolution runs, this is a better choice than selecting the 1D grid 
spacing to be equal to the 3D grid spacing.

@@

%==========================================================================
% Performance
%==========================================================================
\section{Performance}\label{sec:Performance}

strong scaling plot

space filling curve?

OMP vs. MPI?

breakdown by component?

>>>>

We desire to specify the mass of the white dwarf, as well as its
temperature and composition. To start the integration off, we
therefore need to guess at a central density.  We then do a secant
iteration over the entire integration procedure to find the central
density needed to yield the desired total mass.  The grid spacing is
$\Delta x = 6.25\ \text{km}$. We chose this value because no simulation
we perform is likely to exceed this grid resolution inside the stars 
themselves; for our normal domain size (see below), this corresponds to 
three jumps in refinement by a factor of four. We find that for low 
resolution runs, this is a better choice than selecting the 1D grid 
spacing to be equal to the 3D grid spacing.

@@

%==========================================================================
% Performance
%==========================================================================
\section{Parallel Strategy and Performance}\label{sec:Performance}

CASTRO is designed to be deployed on high-performance computing systems using 
many thousands of processors simultaneously. It is worth briefly examining 
our strategy for parallelizing the problem over many computational nodes 
and our performance in situations similar to production science simulations. 
This is especially true because some aspects of our approach to parallelism 
have changed since the first CASTRO paper, and we have obtained improved 
performance in certain settings.

The BoxLib framework that CASTRO is based on splits each AMR level into a number 
of boxes that collectively span the level. These boxes are distributed to processors 
through MPI parallelism; each MPI task in general holds multiple boxes and 
an update includes a loop over all the boxes an MPI task owns. The distribution 
obeys a load-balancing algorithm that attempts to equalize the amount of work 
done by each processor. BoxLib contains a number of strategies for distributing 
work in this way, and by default uses a space-filling curve approach with a 
Morton ordering. By experiment we have found that the most efficient load-balancing 
strategy for our problem is actually a simple knapsack algorithm. In this approach, 
the amount of work owned by a processor is proportional to the number of grid cells 
associated with that processor, and the algorithm attempts to ensure that all 
processors have a similar number of total grid cells. We demand an efficiency of 0.9,
meaning that the average workload per processor should be no smaller than 90\% of the 
maximum workload found on any processor. However, we find that in practice the 
performance is largely insensitive to this choice.

The size and shape of grid boxes is an important consideration for efficiency. 
Boxes that are very small suffer from a host of problems, including the larger 
amount of communication required between hydrodynamics solves. Additionally, 
the multigrid solver will be less efficient if the boxes are small because there 
will be fewer available levels for coarsening and performing V-cycles. Conversely, 
boxes that are too large mean that there isn't enough work to go around when we 
have a large number of processors. Good performance is the result of a careful 
balance between these two effects. On the lower end, we require that all boxes 
be a multiple of 16 zones in each dimension; multigrid efficiency sharply decreases 
if this factor is any lower. On the upper end, we select the maximum grid size 
based on the number of processors we use and the total number of cells in the 
simulation. This size will therefore in general vary on different AMR levels. 
Generally we select a value in between 32 and 64 zones per dimension.

We use OpenMP to accelerate the work associated with the boxes owned by each 
MPI task. Originally CASTRO used OpenMP to accelerate individual loops in the 
hydrodynamics routines, such as the piecewise-parabolic edge state reconstruction 
and the conservative flux update. However, there is a significant amount of 
overhead associated with generating a new OpenMP region at each of the many 
different loops in a hydrodynamics algorithm. This makes such a strategy 
sub-optimal for use on many-core processors and GPUs. We have recently switched 
to a tiling approach, where an OpenMP region is generated at the start of 
the hydrodynamics routine and the individual threads separately work on 
different partitions of each box. This results in much less overhead for 
the threading. In general we obtain more efficient simulations than 
could be obtained using MPI only, because there are fewer boxes and thus 
less communication for a given number of processor cores. We are currently 
developing an approach to evaluating the equation of state and nuclear reactions 
(and possibly the entire hydrodynamics algorithm) on GPUs, which will allow us
to take advantage of the significant computational resources embedded in GPUs on 
certain systems.

\begin{figure}
  \centering
  \includegraphics[scale=0.4]{plots/weak_scaling}
  \caption{CASTRO weak scaling test, performed on the Blue Waters machine at 
    NCSA. Each processor had a fixed amount of work, and we increased the 
    number of simulation zones in concert with the number of processors. The 
    solid curve represents perfect weak scaling, while the blue circles show 
    CASTRO's performance at each processor count. The vertical axis measures 
    the median time per timestep, normalized to this value for the smallest 
    processor count.\label{fig:weak_scaling}}
\end{figure}

To examine the parallel performance of CASTRO, we performed both strong scaling 
and weak scaling tests on the Blue Waters machine at the National Center for 
Supercomputing Applications. For the weak scaling test, whose results are 
shown in Figure \ref{fig:weak_scaling}, we ran a uniform grid binary white 
dwarf simulation for resolutions of $128^3$ zones through $2048^3$ zones. The 
number of processors was scaled with the number of zones so that each processor 
had the same amount of work; the smallest test used 8 processors and the largest 
used 32,768 (note that the number of processor cores on a Blue Waters node is twice 
the number of floating point units on that node). The test was run for 10 timesteps,
with each timestep including two Poisson solves and a hydrodynamics update (though 
for a uniform grid calculation we generally do not need to perform any multigrid 
iterations for the first Poisson solve in a timestep, since the density distribution 
has not changed since the end of the last timestep). We disabled plotfile and 
checkpoint writing, as well as calculation of diagnostic information (the latter 
can contribute to a significant fraction of the run time at large processor 
counts if computed every timestep). We computed the median wall time required per 
time step for each simulation, and then normalized this to the median time per 
timestep for the smallest simulation. We find excellent weak scaling through 
4,096 processors. At the largest run, the simulation time required is 
more than 1.5 times the amount required for the smallest simulation. 
This is due entirely to the increased cost of the multigrid Poisson solve 
in each timestep and this cannot be mitigated except by improving 
communication or computation efficiency in the multigrid solver. We observe 
that this weak scaling behavior with Poisson gravity is a significant 
improvement over the results presented in the first CASTRO paper.

\begin{figure}
  \centering
  \includegraphics[scale=0.4]{plots/strong_scaling}
  \caption{CASTRO strong scaling test performed on the Blue Waters machine at
    NCSA. The vertical axis measures the median time per timestep, and the 
    horizontal axis measures the number of processors in the simulation. Data 
    points are normalized to the time per timestep for the smallest number 
    of processors. The blue circles show the data for a simulation with two AMR
    levels (one coarse and one fine), while the red circles show the data 
    for a test with three AMR levels (one coarse and two fine). The fine levels 
    increase the resolution only in the regions around the stars. For each case 
    we draw a solid curve representing perfect strong scaling.\label{fig:strong_scaling}}
\end{figure}

The strong scaling test we perform uses a grid setup similar to what we 
use for well-resolved binary simulations. With a single refined level, 
we have approximately $5 \times 10^7$ zones spread over $\sim 1000$ grids.
On a second refined level, there are $2 \times 10^9$ zones spread over 
$\sim 16000$ grids. We run a scaling test for both cases, with the 
highest processor count in each case chosen so that the number of 
MPI tasks is similar to the number of grids. There are no relative gains 
to be achieved from further parallelism. The results are found in 
\ref{fig:strong_scaling}. We find excellent scaling for low numbers 
of processors, though the long wall time per timestep required for the 
smallest processor counts makes such a simulation infeasible in practice.
The best tradeoff between reasonable wall time and parallel efficiency 
seems to be when there are approximately 2--4 grids per processor. The 
scaling behavior worsens at the highest processor counts, but this is 
an expected consequence of processors becoming work-starved. In general 
we find very good strong scaling behavior in the regime we are presently 
interested in, simulating the early phases of a simulation at moderate 
resolution. The strong scaling behavior is acceptable, though not perfect, 
at very large processor counts when self-gravity is considered.
